---
title: "Covid91 vaccine study Final2022"
author: "Tracy Alset"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
options(digits=1)
```

```{r}
Males<-filter(FinalData, sex=="M")
Females<- filter(FinalData, sex=="F")
LQBTQ<- filter(FinalData, LGBTQ=="gay")
DrugUsers<- filter(FinalData,  DrugUser=="yes")
```



# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

### Descriptive Results.

## Graphical Results





```{r}
dd2 <- Males %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

Based on the count graph it is looks like the vaccine is working for males because there were less infections in the vaccine group than the placebo group (from the covid 19 bars).
And based on the percentage graph, the percentage of the males who took the placebo and had covid is around 66% which is higher than those who took the vaccine which is around 30%. So far, the vaccine seems to be working for males.



## Numerical Results.

```{r}
table1 <- xtabs(~treatment + infected, data=Males)
rowPerc(table1)
colPerc(table1)
```

Looking at the first table, it looks like 3% out of those who took the placebo had covid, and 1% of those who took the vaccine got covid. This information doesn't really show the efficacy of  the vaccine. But looking at the second table, The majority of those who had covid are the ones who took the placebo( 68%) while it's less infections with the vaccines group ( 32%). So far the vaccine seems to be effective.



### Inferential Results.

```{r}
chisq.test(table1)
chisqtestGC(table1)
malefisher<-fisher.test(table1)
```

Looking at the p-value ( which is very low and it equals 2.645e-11 and it's less than 0.05), we reject  the Null hypothesis and this tells us that there is an actual statistical difference between taking the placebo or the vaccine for men, and this proves that the vaccine is effective for males. 


### Efficacy of vaccine for Males.

```{r}
EfficacyMales<- ((malefisher$estimate/(malefisher$estimate+1))*100)
maleFisherLow<- malefisher$conf.int[1]
maleFisherHigh<- malefisher$conf.int[2]
malefisherlow<- ((maleFisherLow/(maleFisherLow+1))*100)
malefisherhigh<- ((maleFisherHigh/(maleFisherHigh+1))*100)
```

The efficacy of the vaccine among males is `r (EfficacyMales)`%

The 95% confidence interval is from `r (malefisherlow)`% to `r (malefisherhigh)`%

## Females
### Descriptive Results.

## Graphical Results

```{r}
dd2 <- Females %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


Based on the count graph, looking at the covid part of the graph, it looks like there were less infections in the female group that took the vaccine than the female group that took the placebo.
As we move on to the percentage graph, we can see that the percentage of infection in the placebo group which is around 60% is higher than the percentage of infection in the vaccine group which is around 40%. This means that based on our graphical results , so far it looks like the vaccine is effective for females.


## Numerical Results.

```{r}
table2 <- xtabs(~treatment + infected, data=Females)
rowPerc(table2)
colPerc(table2)
```

Looking at the first table , it doesn't look like the data provided helps us to find anything about the effectiveness of the vaccine for women, but based on the information provided by the second table which gives us the percentage of those who took the vaccine or the placebo from the group that had covid, our observations tell us that 59% of those who had covid took the placebo , while 41% of them took the vaccine and this percentage is lower than the one for the placebo, this demonstrates so far that the vaccine is working for females.  

### Inferential Results.

```{r}
chisq.test(table2)
chisqtestGC(table2)
femalefisher<-fisher.test(table2)
```

Looking at the p-value ( which is very low and it equals  0.001 and it's less than 0.05),we reject the Null hypothesis and this tells us that there is an actual statistical difference between taking the placebo or the vaccine for females, and this proves that the vaccine is effective for females.



### Efficacy of vaccine for Females.

```{r}
EfficacyFemales<- ((femalefisher$estimate/(femalefisher$estimate+1))*100)
femaleFisherLow<- femalefisher$conf.int[1]
femaleFisherHigh<- femalefisher$conf.int[2]
femalefisherlow<- ((femaleFisherLow/(femaleFisherLow+1))*100)
femalefisherhigh<- ((femaleFisherHigh/(femaleFisherHigh+1))*100)
```

The efficacy of the vaccine among females is `r (EfficacyFemales)`%

The 95% confidence interval is from `r (femalefisherlow)`% to `r (femalefisherhigh)`%


## LGBTQ

### Descriptive Results.

## Graphical Results




## Numerical Results.



### Inferential Results.



### Efficacy of vaccine for Gays.


## Druggies

### Descriptive Results.

## Graphical Results




## Numerical Results.



### Inferential Results.



### Efficacy of vaccine for Drug users..






